---
title: "Untitled"
author: "Zain"
date: "2025-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}

library(tidyverse)
library(here)

```



```{r}


# Define function to process each dataset
process_survey_data <- function(month, year) {
  
  # Construct file path
  file_path <- here("data", paste0(tolower(month), "_", year, ".csv"))
  
  # Check if file exists
  if (!file.exists(file_path)) {
    warning(paste("File not found:", file_path))
    return(NULL)
  }
  
  # Read the dataset
  df <- read_csv(file_path)
  
  # Select relevant variables (make sure these column names exactly match what's in your CSV files)
  df <- df %>% 
    select(weight,
           "Q44 Over the next 12 months, how likely will you be able to save any money? (probability out of 100)",
           "Q1 How has the financial situation of your household changed over the last 12 months?",
           "Q3 How do you think the general economic situation in this country has changed over the last 12 months?",
           "Q4 How do you expect the general economic situation in this country to develop over the next 12 months?",
           "Q2 How do you expect the financial position of your household to change over the next 12 months?",
           "Q42 Compared to the last 12 months, do you expect to spend more or less money on major purchases such as furniture and electrical goods?")
  
  # Dynamic renaming using paste()
  df <- df %>% rename(
    !!paste0("household_last_12_", tolower(month), "_", year) := "Q1 How has the financial situation of your household changed over the last 12 months?",
    !!paste0("general_economic_last_12_", tolower(month), "_", year) := "Q3 How do you think the general economic situation in this country has changed over the last 12 months?",
    !!paste0("general_economic_next_12_", tolower(month), "_", year) := "Q4 How do you expect the general economic situation in this country to develop over the next 12 months?",
    !!paste0("household_next_12_", tolower(month), "_", year) := "Q2 How do you expect the financial position of your household to change over the next 12 months?",
    !!paste0("major_purchases_", tolower(month), "_", year) := "Q42 Compared to the last 12 months, do you expect to spend more or less money on major purchases such as furniture and electrical goods?",
    !!paste0("save_money_", tolower(month), "_", year) := "Q44 Over the next 12 months, how likely will you be able to save any money? (probability out of 100)"
  )
  
  # Convert save_money to numeric (removing % sign)
  save_money_col <- paste0("save_money_", tolower(month), "_", year)
  df <- df %>%
    mutate(
      !!save_money_col := as.numeric(gsub("%", "", .[[save_money_col]])),
      weight = as.numeric(weight)
    )
  
  # Define weighted net score function
  weighted_net_score <- function(data, var, weight_col) {
    result <- data %>%
      filter(!is.na(.data[[var]])) %>%
      mutate(
        positive_weighted = if_else(.data[[var]] %in% c("A lot better", "A little better", "Spend more money"), .data[[weight_col]], 0),
        negative_weighted = if_else(.data[[var]] %in% c("A lot worse", "A little worse", "Spend less money"), .data[[weight_col]], 0)
      ) %>%
      summarise(
        net_score = ifelse(sum(positive_weighted) + sum(negative_weighted) == 0, NA_real_,
                          round((sum(positive_weighted) - sum(negative_weighted)) / 
                                (sum(positive_weighted) + sum(negative_weighted)) * 100))
      ) %>%
      pull(net_score)
    
    if (length(result) == 0) result <- NA_real_
    return(result)
  }
  
  # Create variable names
  household_last_12_col <- paste0("household_last_12_", tolower(month), "_", year)
  household_next_12_col <- paste0("household_next_12_", tolower(month), "_", year)
  general_economic_last_12_col <- paste0("general_economic_last_12_", tolower(month), "_", year)
  general_economic_next_12_col <- paste0("general_economic_next_12_", tolower(month), "_", year)
  major_purchases_col <- paste0("major_purchases_", tolower(month), "_", year)
  
  # Calculate metrics
  household_last_12_net <- weighted_net_score(df, household_last_12_col, "weight")
  household_next_12_net <- weighted_net_score(df, household_next_12_col, "weight")
  general_economic_last_12_net <- weighted_net_score(df, general_economic_last_12_col, "weight")
  general_economic_next_12_net <- weighted_net_score(df, general_economic_next_12_col, "weight")
  major_purchases_net <- weighted_net_score(df, major_purchases_col, "weight")
  
  # Calculate save_money_net
  save_money_net <- df %>%
    filter(!is.na(.data[[save_money_col]])) %>%
    mutate(
      positive_weighted = if_else(.data[[save_money_col]] >= 60, weight, 0),
      negative_weighted = if_else(.data[[save_money_col]] <= 40, weight, 0)
    ) %>%
    summarise(
      net_score = ifelse(sum(positive_weighted) + sum(negative_weighted) == 0, NA_real_,
                       round((sum(positive_weighted) - sum(negative_weighted)) /
                             (sum(positive_weighted) + sum(negative_weighted)) * 100))
    ) %>%
    pull(net_score)
  
  # Ensure it returns a numeric value and not NULL
  if (length(save_money_net) == 0) save_money_net <- NA_real_
  
  # Calculate overall index
  all_indices <- c(household_last_12_net, household_next_12_net, 
                   general_economic_last_12_net, general_economic_next_12_net, 
                   major_purchases_net)
  overall_index <- round(mean(all_indices, na.rm = TRUE))
  
  # Create a single row with month label and all indices
  result_row <- tibble(
    month_year = paste(month, year),
    household_last_12_net = household_last_12_net,
    household_next_12_net = household_next_12_net,
    general_economic_last_12_net = general_economic_last_12_net,
    general_economic_next_12_net = general_economic_next_12_net,
    major_purchases_net = major_purchases_net,
    save_money_net = save_money_net,
    overall_index = overall_index
  )
  
  return(result_row)
}

# Define list of months and years
months <- c("June", "July", "August", "September", "October", "November", "December", "January", "February")
years <- c(rep("2024", 7), rep("2025", 2))  # Assigning correct years for each month

# Create a data frame to store the results with months as rows and metrics as columns
results_df <- tibble()

# Process each month and add to the results dataframe
for (i in seq_along(months)) {
  month <- months[i]
  year <- years[i]
  message("Processing ", month, " ", year)
  
  result_row <- process_survey_data(month, year)
  if (!is.null(result_row)) {
    results_df <- bind_rows(results_df, result_row)
  }
}

# Add a sortable date column
results_df <- results_df %>%
  mutate(
    month_name = str_extract(month_year, "^\\w+"),
    month_num = match(tolower(month_name), tolower(month.name)),
    year = as.numeric(str_extract(month_year, "\\d{4}$")),
    sort_date = as.Date(paste(year, month_num, "01", sep = "-"))
  ) %>%
  arrange(sort_date) %>%
  select(-month_name, -month_num, -year, -sort_date)

# View the results
print(results_df)

# Save results
# write.csv(results_df, here("results", "monthly_indices_table.csv"), row.names = FALSE)