---
title: "Untitled"
author: "Zain"
date: "2025-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}

library(tidyverse)
library(here)

```



```{r}


# Define function to process each dataset
process_survey_data <- function(month, year) {
  # Construct file path
  file_path <- here("data", paste0(tolower(month), "_", year, ".csv"))
  
  # Check if file exists
  if (!file.exists(file_path)) {
    warning(paste("File not found:", file_path))
    return(NULL)
  }
  
  df <- read_csv(file_path)
  
  
  df <- df %>% 
  select(weight,
         starts_with("Q43"), 
         "Q1 How has the financial situation of your household changed over the last 12 months?",
         "Q3 How do you think the general economic situation in this country has changed over the last 12 months?",
         "Q4 How do you expect the general economic situation in this country to develop over the next 12 months?",
         "Q2 How do you expect the financial position of your household to change over the next 12 months?",
         "Q41 In view of the general economic situation, do you think now is the right time for people to make major purchases such as furniture or electrical goods?")
  
  
  q43_col_name <- names(df)[grep("^Q43", names(df))]

  
  df <- df %>% rename(
  !!paste0("household_last_12_", tolower(month), "_", year) := "Q1 How has the financial situation of your household changed over the last 12 months?",
  !!paste0("general_economic_last_12_", tolower(month), "_", year) := "Q3 How do you think the general economic situation in this country has changed over the last 12 months?",
  !!paste0("general_economic_next_12_", tolower(month), "_", year) := "Q4 How do you expect the general economic situation in this country to develop over the next 12 months?",
  !!paste0("household_next_12_", tolower(month), "_", year) := "Q2 How do you expect the financial position of your household to change over the next 12 months?",
  !!paste0("major_purchases_", tolower(month), "_", year) := "Q41 In view of the general economic situation, do you think now is the right time for people to make major purchases such as furniture or electrical goods?",
   !!paste0("save_money_", tolower(month), "_", year) := q43_col_name
)

  
  
  # Get column names
  household_last_12_col <- paste0("household_last_12_", tolower(month), "_", year)
  household_next_12_col <- paste0("household_next_12_", tolower(month), "_", year)
  general_economic_last_12_col <- paste0("general_economic_last_12_", tolower(month), "_", year)
  general_economic_next_12_col <- paste0("general_economic_next_12_", tolower(month), "_", year)
  major_purchases_col <- paste0("major_purchases_", tolower(month), "_", year)
  save_money_col <- paste0("save_money_", tolower(month), "_", year)  
  
  # Calculate each index
  household_last_12_net <- weighted_net_score(df, household_last_12_col, "weight")
  household_next_12_net <- weighted_net_score(df, household_next_12_col, "weight")
  general_economic_last_12_net <- weighted_net_score(df, general_economic_last_12_col, "weight")
  general_economic_next_12_net <- weighted_net_score(df, general_economic_next_12_col, "weight")
  major_purchases_net <- weighted_net_score(df, major_purchases_col, "weight")
  save_money_net <- weighted_net_score(df, save_money_col, "weight")  
  
  # Calculate overall index
  all_indices <- c(household_last_12_net, household_next_12_net, 
                   general_economic_last_12_net, general_economic_next_12_net, 
                   major_purchases_net, save_money_net)
  overall_index <- round(mean(all_indices, na.rm = TRUE))
  
  # Create a single row with month label and all indices
  result_row <- tibble(
    month_year = paste(month, year),
    household_last_12_net = household_last_12_net,
    household_next_12_net = household_next_12_net,
    general_economic_last_12_net = general_economic_last_12_net,
    general_economic_next_12_net = general_economic_next_12_net,
    major_purchases_net = major_purchases_net,
    save_money_net = save_money_net,
    overall_index = overall_index
  )
  
  return(result_row)
}

# Define the weighted_net_score function OUTSIDE the process_survey_data function
weighted_net_score <- function(data, var, weight_col) {
  result <- data %>%
    filter(!is.na(.data[[var]])) %>%
    mutate(
      positive_weighted = if_else(.data[[var]] %in% c("A lot better", "A little better", "Right time", 
                                                     "a very good time to save.", "a fairly good time to save."), 
                                 .data[[weight_col]], 0),
      negative_weighted = if_else(.data[[var]] %in% c("A lot worse", "A little worse", "Wrong time", 
                                                     "a fairly bad time to save.", "a very bad time to save."), 
                                 .data[[weight_col]], 0)
    ) %>%
    summarise(
      total_weight = sum(.data[[weight_col]]),
      pct_positive_raw = sum(positive_weighted) / total_weight * 100,
      pct_negative_raw = sum(negative_weighted) / total_weight * 100,
      pct_positive = round(pct_positive_raw),  
      pct_negative = round(pct_negative_raw),  
      net_score = pct_positive - pct_negative 
    )
  
  return(result$net_score)
}

# Define list of months and years
months <- c("June", "July", "August", "September", "October", "November", "December", "January", "February")
years <- c(rep("2024", 7), rep("2025", 2))  # Assigning correct years for each month

results_df <- tibble()

for (i in seq_along(months)) {
  month <- months[i]
  year <- years[i]
  message("Processing ", month, " ", year)
  
  result_row <- process_survey_data(month, year)
  if (!is.null(result_row)) {
    results_df <- bind_rows(results_df, result_row)
  }
}

# Add a sortable date column
results_df <- results_df %>%
  mutate(
    month_name = str_extract(month_year, "^\\w+"),
    month_num = match(tolower(month_name), tolower(month.name)),
    year = as.numeric(str_extract(month_year, "\\d{4}$")),
    sort_date = as.Date(paste(year, month_num, "01", sep = "-"))
  ) %>%
  arrange(sort_date) %>%
  select(-month_name, -month_num, -year, -sort_date)


print(results_df)

# Save results

 write.csv(results_df, here("output", "positive_negative.csv"), row.names = FALSE)